FoRTE baseline

### 2019-01-30 analysis/paper/paper.Rmd
PAPER: Sort ED modules based on paper priority

(NOTE: Not a real section; just brainstorming notes for now)

@@

- Branches (`IBRANCH_THERMO`)
        - (0) No branches in energy balance or biophysics
        - (1; default) Branches in energy balance, but not biophysics
        - (2) Branches in both
    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?

@@

- Finite canopy radius (`CROWN_MOD`)
        - (0; default): Complete shading
        - (1): Finite canopy radius (Dietze 2008)
    - See current hypotheses

@@

- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Same as above
- Canopy roughness model (`ICANTURB`)
        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
        - (3) Also Massman, but with varying drag and sheltering within canopy
        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
    - Computes `u*`, `T*`, etc.
        - (1) BRAMS default, based on Louis (1979 TODO). 
        - (2) Oncley and Dudhia (1995 TODO), based on MM5
        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
- Ground to canopy conductance (`IED_GRNDVAP`)
        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
        - (5) Combination of (1) (alpha) and (2) (soil resistance)
    - No idea. Again, need to do some micro-met reading.

@@

- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)

@@

        - (0, default) No treefall disturbance
        - (> 0) Disturbance rate, in 1/years. Will create new patches.
        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
- Growth respiration (`GROWTH_RESP_SCHEME`)
        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass. No effect on overall C budget, but greater within-ecosystem flux resolution.
    - These should produce identical results, and if they don't, that means there is a bug!
    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
    - Note that this whole system is a _hack_, because there is no real "storage respiration" process. In PEcAn, we hard-code the relevant parameters to zero. An interesting question is whether we may be able to get better results with this hack enabled -- i.e. if it compensates a problem in some other process.
- Other configs that ED has, but that we will probably ignore:
    - Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs. Worth checking.
    - Grass (`IGRASS`) -- I'm not running with any grass PFTs.
    - Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
    - Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
    - Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
    - Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
        - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
    - Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances. But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.

>>>>

NOTE: Not a real section; just brainstorming notes for now
Details about these sub-models will probably makes their way into the methods and/or appendices.

### High priority

These are switches that I understand (more-or-less), and that I think will be most relevant to FoRTe.

- Finite canopy radius (`CROWN_MOD`)
    - Options:
        - (0; default): Complete shading
        - (1): Finite canopy radius (Dietze 2008)
    - See current hypotheses
- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
    - Options:
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
    - Options:
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Same as above

### Medium priority

These are switches that may be relevant to FoRTE, but may be harder to pull off.

@@

@@

@@

@@

@@

### Lower priority

These are processes that I think will not have huge impacts, are farther-removed from the FoRTE project goals, or are things I simply don't understand very well (particularly all the micro-meteorology stuff).

- Branches (`IBRANCH_THERMO`)
    - Options:
        - (0) No branches in energy balance or biophysics
        - (1; default) Branches in energy balance, but not biophysics
        - (2) Branches in both
    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
- Canopy roughness model (`ICANTURB`)
    - Options:
        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
        - (3) Also Massman, but with varying drag and sheltering within canopy
        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
    - Computes `u*`, `T*`, etc.
    - Options:
        - (1) BRAMS default, based on Louis (1979 TODO). 
        - (2) Oncley and Dudhia (1995 TODO), based on MM5
        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
- Ground to canopy conductance (`IED_GRNDVAP`)
    - Options:
        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
        - (5) Combination of (1) (alpha) and (2) (soil resistance)
    - No idea. Again, need to do some micro-met reading.
- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
    - Options:
        - (0, default) No treefall disturbance
        - (> 0) Disturbance rate, in 1/years. Will create new patches.
        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
- Growth respiration (`GROWTH_RESP_SCHEME`)
    - Options:
        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass. No effect on overall C budget, but greater within-ecosystem flux resolution.
    - These should produce identical results, and if they don't, that means there is a bug!
    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
    - Note that this whole system is a _hack_, because there is no real "storage respiration" process. In PEcAn, we hard-code the relevant parameters to zero. An interesting question is whether we may be able to get better results with this hack enabled -- i.e. if it compensates a problem in some other process.

### Almost-certainly omit

These configurations are almost certainly irrelevant to the goals of this paper or will not apply at UMBS (e.g. grass or tropical species).
Mostly for reference.

- Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs (TODO: Check that this is the case).
- Grass (`IGRASS`) -- I'm not running with any grass PFTs.
- Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
- Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
- Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
- Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
    - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
- Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances. But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.
